Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations

### 2020-08-27 paper.Rnw
edits through 4.2

\kge{Before introducing the visual diagnostics, let us consider a commonly used visualization of a LIME explanation that is shown in Figure \ref{fig:sine-exp-scatter} (left). The plot is created using the \emph{lime} R package and provides a visual representation of a LIME explanation for the prediction of interest indicated in \autoref{fig:sine-plot-data} (when 4 quantile bins are used to simulate the data).} \sout{To demonstrate a LIME explanation, we focus on the misclassified point indicated in Figure \ref{fig:sine-plot-data} by the diamond as the prediction of interest. The left plot in Figure \ref{fig:sine-exp-scatter} shows a visualization of the explanation for the prediction of interest as suggested by Ribeiro et al and implemented in the \emph{lime} R package.}

@@

\kgc{I moved the next sentence here from the previous paragraph.} Based on the plot \hh{on the left of \autoref{fig:sine-exp-scatter}} alone, it is not possible to make an informed assessment of the explanation. \hh{For a further assessment of the explainer model}, we turn to an \emph{explanation scatterplot}: the \emph{explanation scatterplot} is a visual diagnostic for assessing the LIME claims of locality and fidelity for an individual explanation by juxtaposing the complex and explainer models in one plot. The format of an explanation scatterplot depends on the LIME simulation method. We introduce the explanation scatterplot here under the \emph{lime} R package default method in which the data are simulated uniformly from four quantile bins. In this scenario, LIME converts continuous predictor variables to indicator variables identifying whether the variable value falls in the same quantile bin as the prediction of interest or not. The indicator variables are used as the explainer model features.

The explanation scatterplot is built by plotting the LIME simulated data for the top two features identified by the explanation in a scatterplot and coloring these points by the  predictions from the complex model. The size of the points represents the weight assigned by LIME.
\hh{In order to show the LIME results for the observation of interest, }
 lines \hh{are drawn on top of the points. These lines } represent the boundaries of the indicator variables used to fit the explainer model. The color of the lines denote whether LIME indicates that a feature supports or contradicts a class prediction. Appendix \ref{exp-scatter_plus} addresses the explanation scatterplot formats in other LIME simulation scenarios.

\hh{An explanation scatterplot} \sout{for the} \kge{associated with the LIME explanation depicted on the left in \autoref{fig:sine-exp-scatter}} \sout{example of the previous section} \hh{is shown on the right hand side of \autoref{fig:sine-exp-scatter}.} %The points are the simulated values of $x_2$ versus $x_1$ colored by the random forest predictions and sized by the proximity weight used to fit the ridge regression explainer model. The explainer model is represented by the solid horizontal and vertical lines identifying the boundaries of the quantiles that contain the prediction of interest. The color of the lines indicate whether the explainer model coefficient is positive (blue) or negative (red). Thus, the $x_2$ coefficient is positive, so the $x_2$ value of the prediction of interest supports a random forest prediction of `blue', and the $x_1$ coefficient is negative, so the $x_1$ value of the prediction of interest contradicts a random forest prediction of `blue'. The prediction of interest is represented by the diamond shaped point.
By juxtaposing the random forest predictions and the explainer model boundaries, we are able to assess the faithfulness and localness of the explainer model. First consider the claim of localness. The weights remain relatively high outside of the intersection of the two quantile bins suggesting that the LIME explanation is highly influenced by points outside of the bin containing the prediction of interest. While it is still difficult to say whether or not the claim of localness has been violated, it is possible to say that the weights assigned to the simulated data do not agree with the local region assigned by the intersection of the quantile bins.

Now, consider the claim of faithfulness. Within the intersection of the $x_1$ and $x_2$ quantile bins, there are more random forest probabilities above 0.5. This indicates why the magnitude of the coefficient associated with $x_2$ is larger than that of $x_1$. However, the pattern of the random forest predictions in this plot suggest that better explanations for this prediction exist. One example of a better explanation is to say that the prediction of interest received a random forest prediction greater than 0.5 since the observation of interest falls in the region where $x_1$ is less than 2.5 and $x_2$ is greater 4.5. A more localized explanation is to say that the region around the prediction of interest shows that observations with $-0.3 < x_1 <= 2.5$ and $4.8 < x_2 <= 8.75$ are all assigned probabilities greater than 0.5. The procedure implemented by LIME is not flexible enough to capture either of these explanations.

The explanation scatterplot shows issues with a definition of locality and an oversimplification of the random forest \kge{\sout{model}}. LIME does provide an explanation for the prediction of interest, which makes sense based on the quantile bins used. However, the explanation scatterplot identifies that better explanations for the prediction of interest exist. It is difficult to assess linearity from an explanation scatterplot, but a residual plot of the explainer model could be used to check the linearity claim. See Appendix \ref{residual-plot} for the residual plot associated with the explanation considered in Figure \ref{fig:sine-exp-scatter} and shows a violation of the linearity claim.

The \data \ explanation only includes two features. In situations where more than two features are included in a LIME explanation, the explanation scatterplots can be extended to a generalized pairs plot \citep{emerson:2013} that includes all pairwise combinations of features. Generalized pairs plots (and scatterplot matrices in general) have diminishing value as the number of features increase \citep{jensen:2011} \citep{sweller:2011}. Machine learning models are commonly fit using a large number of features, and therefore, a generalized pairs plot of explanation scatterplots for all features would be ineffective. However, when applying LIME, the user selects the number of features to return in the explanation. In the \emph{lime} R package, \citet{pedersen:2020} encourage users to select less than 10 features to include in the explanation. As a result, specifying a small number of features to include in a LIME explanation makes it feasible to use generalized pairs plot of explanation scatterplots.

@@

# Create the conceptual good heatmap
heatmap_plot_good <- ggplot(data = heatmap_data_good,
                            mapping = aes(x = input,
                                          y = case,
                                          fill = feature,
                                          color = feature)) +
  geom_tile() +
  labs(x = "Tuning Parameter",
       y = "Case",
       fill = "Feature",
       color = "Feature",
       title = "Situation 1: \nConsistent Explanations") +
  theme_bw(base_family = "Times", base_size = 12) +
  scale_fill_grey() +
  scale_color_grey() +
  theme(legend.position = "none")

@@

# Create the conceptual bad heatmap
heatmap_plot_bad <- ggplot(data = heatmap_data_bad,
                            mapping = aes(x = input,
                                          y = case,
                                          fill = feature,
                                          color = feature)) +
  geom_tile() +
  labs(x = "Tuning Parameter",
       y = "Case",
       fill = "Feature",
       color = "Feature",
       title = "Situation 2: \nInconsistent Explanations") +
  theme_bw(base_family = "Times", base_size = 12) +
  scale_fill_grey() +
  scale_color_grey()

@@

@
\caption{Hypothetical examples of feature heatmaps in two possible situations. The heatmaps show the top feature chosen for 10 cases across 5 different sets of LIME tuning parameters. The color of the cell indicates the feature chosen by LIME.}
\label{fig:concept-feat-heat}

@@

<<sine-feat-heat, out.width = '0.45\\textwidth', fig.width = 4, fig.height = 4>>=
plot_feature_heatmap(sine_lime_explain$explain,
                order_method = "PCA") +
  theme_bw(base_family = "Times", base_size = 10) +
  theme(legend.position = "bottom",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.background = element_rect(color = "white", fill = "white")) +
  labs(y = "Case")
@
\caption{\emph{Feature Heatmap}. An example feature heatmap of the explanations from the applications of LIME with different tuning parameters to the \data. The cases from the test set are plotted on the y-axis, and the tuning parameter values (simulation methods here) are included on the x-axis. The colors of the tiles indicate the feature selected by LIME for the corresponding case and tuning parameter value. The plot is faceted by the first features selected by LIME (top facet) and the second most important features (bottom facet). The column facets separate the kernel density method from the quantile bin methods. The vertical striping indicates that the LIME explanations are not consistent across tuning parameters.}
\label{fig:sine-feat-heat}
\end{figure}

Explanations produced by LIME are likely to be affected by the choice of tuning parameters. An example of this is shown by Figures \ref{fig:concept-plot-good} and \ref{fig:concept-plot-bad} where the method used to weight the observations influenced the explanation. As of the time of writing this manuscript, we have encountered no recommendations for how to determine which method to use besides for the default settings in \emph{lime} \kgc{(confirm this)}. In order to compare the explanations produced by LIME using different tuning parameters, we visualize an overview the explanations in the \emph{feature heatmap} diagnostic plot.

The feature heatmap uses colors to identify the features selected by LIME across tuning parameters for sets of explanations (referred to as cases here) faceted by the position of feature importance assigned by LIME. That is, for LIME applied with $t$ sets of tuning parameters to $n$ cases to select the $f$ top features, create $f$ heatmaps (one for each of the positions of importance determined by the magnitude of the explainer model coefficients) with the cases on the y-axis, the tuning parameters on the x-axis, and the cells colored by the feature chosen for the corresponding case and tuning parameter. Additional tuning parameters may also be included in the plot via facets. This plot is used assess the locality of the explanations and to compare results from different input options.

Two hypothetical examples of feature heatmaps are included in \autoref{fig:concept-feat-heat}. The plots were created with the assumption that LIME is applied to select the top feature out of $p=8$ features for $n=10$ cases with $t=5$ sets of tuning parameters. Situation 1 shows an example where the features selected are consistent across tuning parameters within a case but vary across cases within a tuning parameter. This is the ideal situation, because the LIME explanations do not depend on the tuning parameters but do depend on the location of the observation in the feature space. Situation 2 shows an example where the features selected vary across tuning parameters within a case but are consistent across cases within a tuning parameter. This situation indicates that the features selected by LIME are dependent on the tuning parameters, and the explanations are not local because the same feature is chosen regardless of the case. In practice, it is expected that the plot will exhibit a combination of these two situations.

\autoref{fig:sine-feat-heat} shows the feature heatmap for the LIME applications to the 100 observations in the \data \ test set. The most important (top facet) and second most important (bottom facet) features are shown in this plot. For the quantile bins, the original features prior to the indicator variable transformation are included since it is obvious that different features would be selected when the sizes of the bins change. This figure shows that for kernel density and 2 quantile bins, LIME selects $x_1$ as the most important feature and $x_2$ as the second most important feature across all cases in the test set. These explanations are not local. There is variability in the features selected by LIME for 3 to 6 quantile bins. For these tuning parameters, there is a mix of horizontal and vertical stripes, which suggests a dependence on tuning parameters. Note that the explanations from 3 and 5 quantile bins include the selection of the random noise variable ($x_3$) as an important variable in many predictions, which should not be the case. The pattern seen in the explanations for 3 to 6 quantile bins may suggest local explanations, but the dependence on tuning parameter makes it unclear which set of explanations to use.

@@

@
\caption{\emph{Assessment metric plot}. The assessment metrics computed on the applications of LIME to the \data \ test set are included in this example assessment metric plot. Each horizontal facet corresponds to one of three metrics: average $R^2$, average fidelity, or MSEE. The LIME tuning parameter (simulation methods in this example) is plotted on the x-axis, and the metric values are shown on the y-axis. The points are colored by rank (within a metric) indicating best to worst (dark to light). The kernel density simulation methods performs well according to all three metrics.}
\label{fig:sine-assess-metric}
\end{figure}

@@

Each metric presented below is computed on a set of LIME explanations obtained using the the same tuning parameters. Here we provide a high level description of the metrics. Notation and formulas for these metrics are included in Appendix \ref{metric-details}.

@@

\autoref{fig:sine-assess-metric} shows these metrics computed for each of the LIME applications to the \data \ test set. The simulation methods are listed on the x-axis. The plot is faceted by metric, and the metric values are plotted on the y-axis. The colors of the points represent the rank of the simulation methods performance based on a particular metric (darker indicates a better metric value and lighter indicates a worse metric value). Higher average $R^2$ values are better, and lower average fidelity and MSEE values are better. This example only includes one tuning parameter: the simulation method. If more than one tuning parameter is considered, the assessment metric plot is extended by adding additional facets or levels to the x-axis.

All three metrics suggest that the kernel density performed well, but the metrics disagree for the quantile bins methods. Average $R^2$ and average fidelity rank the performance of the number of quantile bins the same (2 quantile bins perform the best and 6 quantile bins perform the worst). In fact, these two metrics appear to have an inverse relationship in this example. MSEE provides different ranks of the simulation methods. MSEE suggests that 6 quantile bins perform the best, and 2 quantile bins perform the worst. This is the opposite of average $R^2$ and average fidelity. Since average fidelity and MSEE are similar metrics, it would be expected that they would agree, but that is so in this example. This may be due to the MSEE only taking into account the prediction of interest and not the full simulated dataset. The contradiction between metrics makes it difficult to identify which simulation method explanations to trust.

Recall that \autoref{fig:sine-feat-heat} indicated that the LIME kernel density method selected the same feature across all cases in the test set for both the first and second features. It appears that a global trend may be the best explanation for this example, which may be reasonable considering that we know that both $x_1$ and $x_2$ are the two features that should be the features used by the random forest to distinguish between response categories.

@@

In current practice, forensic firearm examiners evaluate whether two bullets are from the same source (fired from the same gun) or from different sources based on microscopic comparison of the striation patterns engraved on bullets during the firing process (see \autoref{fig:bullet}). The process is based on a visual and therefore subjective assessment of the evidence. The lack of objective evaluation and the associated absence of established error rates has first been criticized by the National Research Council \cite{nrc:2009} and later by the President's Council of Advisors on Science and Technology \cite{pcast:2016}.

In response, \citet{hare:2016} proposed an automated machine learning method for bullet matching to complement a visual inspection by firearm examiners. Based on high-resolution topological scans of land engraved areas \citet{hare:2016} obtain signatures of striations from two bullet lands (\autoref{fig:signatures}). Nine features quantifying the similarity of signatures, such as the cross-correlation function, the distance between signatures, and the number of matching striae, are extracted and used to train a random forest \kge{\sout{model}} (available in the \emph{bulletxtrctr} R package) to determine the probability of a comparison resulting from the same source (matching signatures) or from different sources (non-matching signatures). The model was trained on a set of scans of bullets from the James Hamby Consecutively Rifled Ruger Barrel Study \citep{hamby:2009}, which includes \Sexpr{dim(bullet_train)[1]} land-to-land comparisons.

@@

\includegraphics[width=3.25in]{./figure-static/signatures.png}
\label{fig:signatures}
\caption{Example bullet signatures. The bullet signatures are from the same barrel-land and therefore have very similar patterns. The \citet{hare:2016} random forest uses features that measure the similarity between two such signatures.}
\end{figure}

@@

We apply the random forest \kge{\sout{model}} from \citet{hare:2016} to another set of bullets from the Hamby study with \Sexpr{dim(bullet_test)[1]} rows of land comparisons. We first consider a global visualization of the relationship between the random forest predictions and the model features by creating a parallel coordinate plot of the training data (top facet) and testing data (bottom facets) predictions (\autoref{fig:bullet-plot-rfpcp}). The majority of same source observations with random forest probabilities close to 1 have a clear pattern of corresponding feature values. The known same source observations where the random forest is wrong (does not return probabilities close to 1) have feature values that reflect those of observations where same source is known to be false.

LIME is implemented using the R package \emph{limeaid} to apply LIME multiple times to all test set observations using different tuning parameters: 12 sampling methods (2-6 equally spaced bins, 2-6 quantile bins, kernel density estimation, and normal approximation) and 3 Gower exponents (0.5, 1, and 10). Thus, a total of $12\times 3=36$ different applications of LIME were performed. We specify that each LIME explanation return 3 features and feature selection is performed using the highest weights method (default option).

>>>>

\kge{Before introducing the visual diagnostics, let us consider a commonly used visualization of a LIME explanation that is shown in Figure \ref{fig:sine-exp-scatter} (left). The plot is created using the \emph{lime} R package and provides a visual representation of a LIME explanation for the prediction of interest indicated in \autoref{fig:sine-plot-data}. In particular, the explanation was obtained using 4 quantile bins to simulate the data.} \kgc{I moved the next few sentences here from the sub-section.} In this scenario, LIME converts continuous predictor variables to indicator variables identifying whether the variable value falls in the same quantile bin as the prediction of interest or not. The indicator variables are used as the explainer model features. \sout{To demonstrate a LIME explanation, we focus on the misclassified point indicated in Figure \ref{fig:sine-plot-data} by the diamond as the prediction of interest. The left plot in Figure \ref{fig:sine-exp-scatter} shows a visualization of the explanation for the prediction of interest as suggested by Ribeiro et al and implemented in the \emph{lime} R package.}

@@

\kgc{I moved the next sentence here from the previous paragraph.} Based on the plot \hh{on the left of \autoref{fig:sine-exp-scatter}} alone, it is not possible to make an informed assessment of the explanation. \hh{For a further assessment of the explainer model}, we turn to an \emph{explanation scatterplot}: \sout{the \emph{explanation scatterplot} is} a visual diagnostic for assessing the LIME claims of locality and fidelity for an individual explanation by juxtaposing the complex and explainer models in one plot. The format of an explanation scatterplot depends on the LIME simulation method. We introduce the explanation scatterplot here under the \emph{lime} R package default method \kge{four quantile bins.} \sout{in which the data are simulated uniformly from four quantile bins.}

The explanation scatterplot is built by plotting the LIME simulated data for the top two features identified by the explanation in a scatterplot and coloring these points by the  predictions from the complex model. The \kge{point} size \sout{of the points} represents the weight assigned by LIME.
\hh{In order to show the LIME results for the observation of interest,} lines \hh{are drawn on top of the points. These lines } represent the boundaries of the indicator variables used to fit the explainer model. The \kge{line} color \sout{of the lines} denote\kge{s} whether LIME indicates that a feature supports or contradicts a class prediction. Appendix \ref{exp-scatter_plus} addresses the explanation scatterplot formats in other LIME simulation scenarios.

\hh{An explanation scatterplot} \sout{for the} \kge{corresponding to the LIME explanation depicted on the left in \autoref{fig:sine-exp-scatter}} \sout{example of the previous section} \hh{is shown on the right hand side of \autoref{fig:sine-exp-scatter}.} %The points are the simulated values of $x_2$ versus $x_1$ colored by the random forest predictions and sized by the proximity weight used to fit the ridge regression explainer model. The explainer model is represented by the solid horizontal and vertical lines identifying the boundaries of the quantiles that contain the prediction of interest. The color of the lines indicate whether the explainer model coefficient is positive (blue) or negative (red). Thus, the $x_2$ coefficient is positive, so the $x_2$ value of the prediction of interest supports a random forest prediction of `blue', and the $x_1$ coefficient is negative, so the $x_1$ value of the prediction of interest contradicts a random forest prediction of `blue'. The prediction of interest is represented by the diamond shaped point.
By juxtaposing the random forest predictions and the explainer model boundaries, we are able to assess the faithfulness and localness of the explainer model.

First, consider the claim of localness. The weights \kge{decay relatively slowly} \sout{remain relatively high} outside of the intersection of the two quantile bins suggesting that the LIME explanation is highly influenced by points outside of the bin\kge{s} containing the prediction of interest. \sout{While i} \kge{However,} it is \sout{still} difficult to say \kge{if} \sout{whether or not} the claim of localness has been violated\kge{, because a definition of a local region was not specified. Depending a the region a viewer considers to be local, an argument could be made in favor of or against a violation of localness. The explanation scatterplot raises awareness of the unclear definition of a local region with LIME. Nevertheless,} it is possible to say that the weights assigned to the simulated data do not \kge{capture} \sout{agree with} the local region \kge{identified} \sout{assigned} by the intersection of the quantile bins.

\kgc{I tried rewriting this paragraph to make it less "technical". See what you think.} Now, consider the claim of faithfulness. \kge{It can be said that the majority of the points in the $x_2$ quantile support a prediction of blue, which is captured by the bar supporting a prediction of blue in LIME explanation, and a similar statement can be said about the $x_1$ quantile bin. These statements validate the explanation produced by LIME. However, the explanation scatterplot plot shows that the random forest does a good job of capturing the sine curve classification boundary by creating various sized rectangles with similar probabilities, which the LIME explanation does not pick up on. Thus, LIME does provide an explanation for the prediction, but it is a very poor explanation in terms of faithfulness to the random forest prediction regions.}

\sout{Within the intersection of the $x_1$ and $x_2$ quantile bins, there are more random forest probabilities above 0.5 \kge{than below}. This indicates why the magnitude of the coefficient associated with $x_2$ is larger than that of $x_1$. However, the pattern of the random forest predictions in this plot suggest that better explanations for this prediction exist. One example of a better explanation is to say that the prediction of interest received a random forest prediction greater than 0.5 since the observation of interest falls in the region where $x_1$ is less than 2.5 and $x_2$ is greater 4.5. A more localized explanation is to say that the region around the prediction of interest shows that observations with $-0.3 < x_1 <= 2.5$ and $4.8 < x_2 <= 8.75$ are all assigned probabilities greater than 0.5. The procedure implemented by LIME is not flexible enough to capture either of these explanations.}

\sout{The explanation scatterplot shows issues with a definition of locality and an oversimplification of the random forest \kge{\sout{model}}. LIME does provide an explanation for the prediction of interest, which makes sense based on the quantile bins used. However, the explanation scatterplot identifies that better explanations for the prediction of interest exist.}

It is difficult to assess linearity from \kge{this} \sout{an} explanation scatterplot, but a residual plot of the explainer model could be used to check the linearity claim. See Appendix \ref{residual-plot} for the residual plot associated with the explanation considered in Figure \ref{fig:sine-exp-scatter}, which shows a violation of the linearity claim.

\kge{This example} \sout{The \data \ } explanation only includes two features. In situations where more than two features are included in a LIME explanation, the explanation scatterplots can be extended to a generalized pairs plot \citep{emerson:2013} that includes all pairwise combinations of features. Generalized pairs plots (and scatterplot matrices in general) have diminishing value as the number of features increase \citep{jensen:2011} \citep{sweller:2011}. Machine learning models are commonly fit using a large number of features, and therefore, a generalized pairs plot of explanation scatterplots for all features would be ineffective. However, when applying LIME, the user selects the number of features to return in the explanation. In the \emph{lime} R package, \citet{pedersen:2020} encourage users to select less than 10 features \sout{to include in the explanation}. \kge{As long as a small number of features are returned in the LIME explanation,} \sout{As a result, specifying a small number of features to include in a LIME explanation makes} it \kge{is} feasible to use \kge{a} generalized pairs plot of explanation scatterplots. \kge{An example is shown in Section \ref{bullet-assess-ex}.}

@@

# Create the conceptual good heatmap
heatmap_plot_good <- ggplot(data = heatmap_data_good,
                            mapping = aes(x = input,
                                          y = case,
                                          fill = feature,
                                          color = feature)) +
  geom_tile() +
  labs(x = "Tuning Parameter Value",
       y = "Case",
       fill = "Feature",
       color = "Feature",
       title = "Situation 1: \nConsistent Explanations") +
  theme_bw(base_family = "Times", base_size = 12) +
  scale_fill_grey() +
  scale_color_grey() +
  theme(legend.position = "none")

@@

# Create the conceptual bad heatmap
heatmap_plot_bad <- ggplot(data = heatmap_data_bad,
                            mapping = aes(x = input,
                                          y = case,
                                          fill = feature,
                                          color = feature)) +
  geom_tile() +
  labs(x = "Tuning Parameter Value",
       y = "Case",
       fill = "Feature",
       color = "Feature",
       title = "Situation 2: \nInconsistent Explanations") +
  theme_bw(base_family = "Times", base_size = 12) +
  scale_fill_grey() +
  scale_color_grey()

@@

@
\caption{Hypothetical examples of feature heatmaps in two possible situations. The heatmaps show the top feature chosen for 10 cases across 5 different sets of \sout{LIME} tuning parameters. The color of the cell indicates the feature chosen by LIME. \kge{Situation 1 is the ideal, because the explanations are not dependent on tuning parameter values.}}
\label{fig:concept-feat-heat}

@@

<<sine-feat-heat, out.width = '0.45\\textwidth', fig.width = 4, fig.height = 4>>=
plot_feature_heatmap(sine_lime_explain$explain,
                order_method = "PCA") +
  theme_bw(base_family = "Times", base_size = 10) +
  theme(legend.position = "bottom",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.background = element_rect(color = "white", fill = "white")) +
  labs(y = "Case")
@
\caption{\emph{Feature Heatmap}. An example feature heatmap of the explanations from the applications of LIME with different tuning parameters to the \data \ test set. The cases from the test set are plotted on the y-axis, and the tuning parameter values (simulation methods here) are included on the x-axis. The colors of the tiles indicate the feature selected by LIME for the corresponding case and tuning parameter value. The plot is faceted by the first \kge{and second most important} features selected by LIME (top \kge{and bottom} facet\kge{s, respectively}) \sout{ and the second most important features (bottom facet).} \kge{The vertical facets separate the} \sout{The column facets separate the} kernel density method from the quantile bin methods. The vertical striping indicates that the LIME explanations are not consistent across tuning parameters.}
\label{fig:sine-feat-heat}
\end{figure}

Explanations produced by LIME are likely to be affected by the choice of tuning parameters. \sout{An} \kge{A hypothetical} example of this is shown by Figures \ref{fig:concept-plot-good} and \ref{fig:concept-plot-bad} where the method used to weight the observations influenced the explanation. As of the time of writing this manuscript, we have encountered no recommendations for how to \kge{specify the parameter values} \sout{determine which method to use} besides for the default settings in \emph{lime}. In order to compare the explanations produced by LIME using different tuning parameters \kge{and provide another perspective for assessing localness}, we visualize an overview \kge{of} the explanations \kge{with} \sout{in} the \emph{feature heatmap} diagnostic plot.

The feature heatmap uses colors to identify the features selected by LIME \kge{across multiple predictions (referred to as cases here) and tuning paramter values} \sout{across tuning parameters for sets of explanations (referred to as cases here)} \kge{organized} \sout{faceted} by the \sout{position of} feature importance assigned by LIME. That is, for LIME applied with $t$ sets of tuning parameters to $n$ cases to select the $f$ top features, create $f$ heatmaps (one for each of the positions of importance determined by the magnitude of the explainer model coefficients) with the cases on the y-axis, the tuning parameters on the x-axis, and the cells colored by the feature chosen for the corresponding case and tuning parameter. Additional tuning parameters may also be included in the plot via facets. \sout{This plot is used assess the locality of the explanations and to compare results from different input options.}

Two hypothetical examples of feature heatmaps are included in \autoref{fig:concept-feat-heat}. The plots were created with the assumption that LIME is applied to select the top feature out of $p=\kge{4}$ \sout{8} features for $n=10$ cases with $t=5$ sets of tuning parameters. Situation 1 \kge{is} \sout{shows} an example where the features selected are consistent across tuning parameters within a case but vary across cases within a tuning parameter. This is the ideal situation, because the LIME explanations do not depend on the tuning parameters but do depend on the location of the observation in the feature space. Situation 2 \kge{is} \sout{shows} an example where the features selected vary across tuning parameters within a case but are consistent across cases within a tuning parameter. This situation indicates that the features selected by LIME are dependent on the tuning parameters, and the explanations \kge{may not be} \sout{are not} local, because the same feature is chosen regardless of the case. In practice, it is expected that the plot will exhibit a combination of these two situations.

\autoref{fig:sine-feat-heat} shows \kge{a} \sout{the} feature heatmap for the LIME applications to the 100 observations in the \data \ test set. The most important \kge{and second most important features selected by lime are showin in the top and bottom facets, respectively.} \sout{(top facet) and second most important (bottom facet) features are shown in this plot.} For the quantile bins, the original features prior to the indicator variable transformation are included since it is obvious that different features would be selected when the sizes of the bins change. This figure shows that for kernel density and 2 quantile bins, LIME selects $x_1$ as the most important feature and $x_2$ as the second most important feature across all cases in the test set. These explanations are not local. There is variability in the features selected by LIME for 3 to 6 quantile bins \kge{suggesting more local explanations}. \sout{For these tuning parameters, there is a mix of horizontal and vertical stripes,} \kge{There are signs of vertical striping,} which suggests a dependence on tuning parameters. Note that the explanations from 3 and 5 quantile bins include the selection of the random noise variable ($x_3$) as an important variable in many predictions, which should not be the case. The pattern seen in the explanations for 3 to 6 quantile bins may suggest local explanations, but the dependence on tuning parameter makes it unclear which set of explanations to use.

@@

@
\caption{\emph{Assessment metric plot}. The assessment metrics computed on the applications of LIME to the \data \ test set are included in this example assessment metric plot. Each horizontal facet corresponds to one of three metrics: average $R^2$, average fidelity, or MSEE. The LIME tuning parameter \kge{values} (simulation methods in this example) \kge{are} \sout{is} plotted on the x-axis, and the metric values are shown on the y-axis. The points are colored by rank (within a metric) indicating best to worst (dark to light). The kernel density simulation methods performs well \kge{across} \sout{according to} all three metrics.}
\label{fig:sine-assess-metric}
\end{figure}

@@

Each metric presented below is computed on a \sout{set of} LIME explanations \kge{for a set of predictions} obtained using the \sout{the} same tuning parameters. Here we provide a high level description of the metrics. Notation and formulas for these metrics are included in Appendix \ref{metric-details}.

@@

\autoref{fig:sine-assess-metric} shows \kge{an example assessment metric plot}. \sout{these} \kge{The three} metrics \kge{are} computed for each of the LIME applications to the \data \ test set. The simulation methods are listed on the x-axis. The plot is faceted by metric, and the metric values are plotted on the y-axis. The colors of the points represent the rank of the simulation methods performance based on a particular metric (darker indicates a better metric value and lighter indicates a worse metric value). Higher average $R^2$ values are better, and lower average fidelity and MSEE values are better. This example only includes one tuning parameter: the simulation method. If more than one tuning parameter is considered, the assessment metric plot is extended by adding additional facets or levels to the x-axis.

All three metrics suggest that the kernel density \kge{method} performed well, but the metrics disagree for the quantile bins methods. Average $R^2$ and average fidelity rank the performance of the number of quantile bins the same (2 quantile bins perform the best and 6 quantile bins perform the worst). In fact, these two metrics appear to have \kge{a mirrored} \sout{an inverse} relationship in this example. MSEE provides \kge{almost the exact opposite results with} \sout{different ranks of the simulation methods. MSEE suggests that} 6 quantile bins perform\kge{ing} the best\sout{,} and 2 quantile bins perform\kge{ing} the worst. \sout{This is the opposite of average $R^2$ and average fidelity.} Since average fidelity and MSEE are similar metrics, it would be expected that they would agree, but that is so in this example. This may be due to the MSEE only taking into account the prediction of interest and not the full simulated dataset. The contradiction between metrics makes it difficult to identify which simulation method \sout{explanations} to trust.

Recall that \autoref{fig:sine-feat-heat} indicated that the \sout{LIME} kernel density method selected the same feature across all cases in the test set for both the first and second features. It appears that a global trend may be the best explanation for this example, which may be reasonable considering that we know that both $x_1$ and $x_2$ are the two features that should be the features used by the random forest to distinguish between response categories.

@@

In current practice, forensic firearm examiners evaluate whether two bullets are from the same source (fired from the same gun) or from different sources based on \kge microscopic comparison of the striation patterns engraved on bullets during the firing process (see \autoref{fig:bullet}). The process is based on a visual and therefore subjective assessment of the evidence. The lack of objective evaluation and the associated absence of established error rates has first been criticized by the National Research Council \cite{nrc:2009} and later by the President's Council of Advisors on Science and Technology \cite{pcast:2016}.

In response, \citet{hare:2016} proposed an automated machine learning method for bullet matching to complement a visual inspection by firearm examiners. Based on high-resolution topological scans of land engraved areas\kge{,} \citet{hare:2016} obtain signatures of striations from two bullet lands (\autoref{fig:signatures}). Nine features quantifying the similarity of signatures, such as the cross-correlation function, the distance between signatures, and the number of matching striae, are extracted and used to train a random forest \kge{\sout{model}} (available in the \emph{bulletxtrctr} R package) to determine the probability of a comparison resulting from the same source (matching signatures) or from different sources (non-matching signatures). The model was trained on a set of scans of bullets from the James Hamby Consecutively Rifled Ruger Barrel Study \citep{hamby:2009}, which includes \Sexpr{dim(bullet_train)[1]} land-to-land comparisons.

@@

\includegraphics[width=3.25in]{./figure-static/signatures.png}
\caption{Example bullet signatures. The bullet signatures are from the same barrel \kge{and} land and therefore have very similar patterns. The \citet{hare:2016} random forest \kge{is fit using various} \sout{uses} features that measure the similarity between two such signatures.}
\label{fig:signatures}
\end{figure}

@@

We apply the random forest \kge{\sout{model}} from \citet{hare:2016} to another set of bullets from the Hamby study with \Sexpr{dim(bullet_test)[1]} rows of land comparisons. We first consider a global visualization of the relationship between the random forest predictions and the model features \kge{with} \sout{by creating} a parallel coordinate plot of the training data (top facet) and testing data (bottom facets) predictions (\autoref{fig:bullet-plot-rfpcp}). The majority of same source observations with random forest probabilities close to 1 have a clear pattern of corresponding feature values. The known same source observations where the random forest is wrong (does not return probabilities close to 1) have feature values that reflect those of observations where same source is known to be false.

LIME \sout{is implemented using the R package \emph{limeaid} to apply LIME} \kge{is applied} \sout{multiple times} to all test set observations using different tuning parameters: 12 sampling methods (2-6 equally spaced bins, 2-6 quantile bins, kernel density estimation, and normal approximation) and 3 Gower exponents (0.5, 1, and 10). Thus, a total of $12\times 3=36$ different applications of LIME were performed. We specify that each LIME explanation return 3 features and feature selection is performed \kge{by selecting the features with the highest weights in a ridge regression model} \sout{using the highest weights method} (default option). \kgc{XXX THIS IS WHERE I LEFT OFF XXX}
